Energy Clustering

### 2018-06-04 energy_clustering/pami18/old_version/energy_clustering.tex
energy pami

\documentclass[aps,preprint,nofootinbib,floatfix]{revtex4-1}

@@

%\pretolerance=10000
%\tolerance=2000 
%\emergencystretch=10pt

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}{Example}

@@

\begin{document}

\title{Energy Clustering}

\author{Guilherme Fran\c ca}
\email{guifranca@gmail.com} 
\author{Joshua T. Vogelstein}
\email{jovo@jhu.edu}
\affiliation{Johns Hopkins University}

\begin{abstract}
Energy statistics was proposed by Sz\' ekely in the 80's inspired by the 
Newtonian gravitational potential from classical mechanics, and it provides 
a 
hypothesis test for equality of distributions. 
It was further generalized from Euclidean spaces to metric spaces of 
strong negative type, and more recently, a connection with reproducing 
kernel Hilbert spaces (RKHS) was established. 
Here we consider the clustering problem from an 
energy statistics theory perspective, 
providing a precise mathematical formulation 
yielding a quadratically constrained 
quadratic program (QCQP) in the associated RKHS, thus establishing the
connection with kernel methods. We show that this QCQP
is equivalent to kernel $k$-means optimization problem once the kernel
is fixed.
These results imply a first principles derivation of kernel $k$-means 
from energy statistics.
However, energy statistics fixes a family of standard kernels.
Furthermore, we also consider a weighted version of energy statistics, 
making connection to graph partitioning problems.
To find local optimizers of such QCQP we propose an iterative algorithm based 
on Hartigan's method, which in this case has the same computational cost 
as kernel $k$-means algorithm, based on Lloyd's heuristic, but usually 
with better clustering quality. 
We provide carefully designed numerical experiments showing the superiority 
of the proposed method compared to kernel $k$-means, spectral clustering,
standard $k$-means, and Gaussian mixture models in a variety of settings.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Energy statistics \cite{Szkely2013}
is based on a 
notion of statistical potential energy between probability distributions,
in close analogy to Newton's gravitational potential in classical mechanics.
It provides a model-free hypothesis test for equality of 
distributions which is achieved 
under minimum energy. When probability distributions are different the 
statistical potential energy diverges as sample size increases, while tends 
to a nondegenerate limit distribution when probability
distributions are equal. 
Energy statistics has been applied to several goodness-of-fit 
hypothesis tests, multi-sample tests of equality of distributions, 
analysis of variance \cite{RizzoVariance}, nonlinear dependence tests through
distance covariance and distance correlation, which generalizes the Pearson
correlation coefficient, and hierarchical clustering \cite{RizzoClustering} 
by extending Ward's method of minimum variance. Moreover, 
in Euclidean spaces, an application of 
energy statistics to clustering
was already proposed \cite{Kgroups}. 
We refer the reader to \cite{Szkely2013}, 
and references therein, for an overview of energy statistics theory and 
its applications.

In its original formulation, energy statistics has a compact representation
in terms of expectations of pairwise Euclidean distances, providing
straightforward empirical estimates. 
The notion of distance covariance was further generalized from Euclidean 
spaces to metric spaces of strong negative type \cite{Lyons}. Furthermore, 
the missing link between energy distance based tests and kernel 
based tests has 
been recently resolved \cite{Sejdinovic2013}, 
establishing an equivalence between generalized energy distances to maximum
mean discrepancies (MMD), which are distances between embeddings of 
distributions in reproducing kernel Hilbert spaces (RKHS).
This equivalence immediately relates energy statistics to
kernel methods often used in machine learning, and form the basis 
of our approach.

@@

Although clustering from energy statistics, in Euclidean spaces,
was considered in \cite{Kgroups}, 
the precise optimization problem behind this approach
remains elusive, as well as the connection with kernel methods.
The main theoretical contribution of this paper is to fill this gap.
Since the statistical potential energy is minimum when
distributions are equal, the principle behind clustering is to maximize 
the statistical energy,  enforcing probability distributions associated to 
each cluster to be different from one another. We provide a precise 
mathematical formulation to this statement, leading to a quadratically 
constrained quadratic program (QCQP) in the associated RKHS. This immediately
establishes the connection between energy statistics based clustering,
or \emph{energy clustering} for short, with kernel methods. Moreover,
our formulation holds for general semimetric spaces of negative type.
We also show that such QCQP is equivalent to kernel $k$-means 
optimization problem, however, the kernel is fixed by 
energy statistics.
The equivalence between kernel $k$-means, spectral clustering, and graph 
partitioning problems is well-known \cite{Dhillon,Dhillon2}. We further
demonstrate 
how these relations arise from a weighted version of energy statistics.

Our main 
algorithmic contribution is to use Hartigan's method \cite{Hartigan} to 
find local solutions of the above mentioned QCQP, which is NP-hard in general.
Hartigan's method was also used in \cite{Kgroups}, but without
any connection to kernels. More importantly, the
advantages
of Hartigan's over Lloyd's method was already demonstrated 
in some simple settings
\cite{Telgarsky,Slonin}, but apparently this method did not receive 
the deserved attention. To the 
best of our knowledge, Hartigan's method was not previously 
employed together with kernel methods. 
We provide a fully kernel based Hartigan's algorithm for clustering,
where the kernel is fixed by energy statistics. 
We make clear the advantages of this proposal versus 
Lloyd's method, which kernel $k$-means is based upon and will also be used 
to solve our QCQP. We show that both algorithms  have the same
time complexity, but Hartigan's method in kernel spaces offer several
advantages. Furthermore, in the examples considered in this paper, it 
also provides superior performance compared spectral clustering, which
is more expensive and in fact solves a relaxed version of our QCQP.

Our numerical results provide compelling evidence that 
Hartigan's method applied to energy clustering 
is more accurate and robust than kernel $k$-means algorithm.
Furthermore, our experiments illustrate 
the flexibility of energy clustering, 
showing that it is able to perform accurately on data coming from 
very different distributions, contrary to $k$-means and GMM for instance.
More specifically, the proposed method performs 
closely to $k$-means and GMM on normally distributed data, however,
it is significantly better on data that 
is not normally distributed. 
Its superiority in high dimensions
is striking, being more accurate than $k$-means and GMM 
even on Gaussian settings.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background on Energy Statistics and RKHS}
\label{sec:background}

>>>>

%% bare_jrnl_compsoc.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% Computer Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[10pt,journal,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

@@

%\pretolerance=10000
%\tolerance=2000 
%\emergencystretch=10pt
\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}{Example}

@@

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Energy Clustering}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Guilherme~Fran\c ca,~Maria Rizzo~and~Joshua~Vogelstein
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem G. Fran\c ca is
with the Center for Imaging Science, Johns Hopkins
University; J. Vogelstein is with the Center for Imaging Science,
Department of Biomedical Engineering and Institute for Computational
Medicine, Johns Hopkins University. \protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: guifranca@jhu.edu; jovo@jhu.edu
\IEEEcompsocthanksitem M. Rizzo is with the Department of Mathematics
and Statistics, Bowling Green State University. \protect\\
E-mail: mrizzo@bgsu.edu}% <-this % stops an unwanted space
\thanks{Manuscript received March 15, 2018; revised March 15, 2018.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Energy Clustering}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Energy statistics was proposed by Sz\' ekely in the 80's inspired by the
Newtonian gravitational potential from classical mechanics and it provides
a hypothesis test for equality of distributions.
It was further generalized from Euclidean spaces to metric spaces of
strong negative type, and more recently, a connection with reproducing
kernel Hilbert spaces (RKHS) was established.
In this paper, we consider the problem of clustering data 
from the perspective of energy statistics theory. 
We provide a precise mathematical formulation 
yielding a quadratically constrained
quadratic program (QCQP) in the associated RKHS, thus establishing the
connection with kernel methods. 
%We show that this QCQP
%is equivalent to kernel $k$-means optimization problem once the kernel
%is fixed.
%These results imply a first principles derivation of kernel $k$-means
%from energy statistics.
%However, energy statistics fixes a family of standard kernels.
Moreover, we also propose a weighted version of energy statistics applied
to clustering, making connection to graph partitioning problems.
To find local solutions of such QCQP, we propose an iterative algorithm based
on Hartigan's method, which in this case has the same computational cost
as kernel $k$-means algorithm, which is on Lloyd's heuristic, but usually
with better clustering quality.
We provide carefully designed numerical experiments showing the superiority
of the proposed method compared to kernel $k$-means, spectral clustering,
standard $k$-means, and Gaussian mixture models in a variety of settings,
specially in high dimensions. We also 
employ the proposed method to an important
real dataset describing protein expressions of neural synapses.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Clustering, Energy Statistics, Kernel Methods.
\end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.

\IEEEPARstart{E}{nergy Statistics} \cite{Szkely2013}
is based on a 
notion of statistical potential energy between probability distributions,
in close analogy to Newton's gravitational potential in classical mechanics.
It provides a model-free hypothesis test for equality of 
distributions which is achieved 
under minimum energy. When probability distributions are different the 
``statistical potential energy'' diverges as sample size increases, 
while tends 
to a nondegenerate limit distribution when probability
distributions are equal. 
Energy statistics has been applied to several goodness-of-fit 
hypothesis tests, multi-sample tests of equality of distributions, 
analysis of variance \cite{RizzoVariance}, nonlinear dependence tests through
distance covariance and distance correlation, which generalizes the Pearson
correlation coefficient, and hierarchical clustering \cite{RizzoClustering} 
by extending Ward's method of minimum variance. Moreover, 
in Euclidean spaces, an application of 
energy statistics to clustering
was already proposed \cite{Kgroups}. 
We refer the reader to \cite{Szkely2013}, 
and references therein, for an overview of energy statistics theory and 
its applications.

In its original formulation, energy statistics has a compact representation
in terms of expectations of pairwise Euclidean distances, providing
straightforward empirical estimates. 
The notion of distance covariance was further generalized from Euclidean 
spaces to metric spaces of strong negative type \cite{Lyons}. Furthermore, 
the link between energy distance based tests and kernel 
based tests has 
been recently established \cite{Sejdinovic2013} 
through an equivalence between generalized energy distances to maximum
mean discrepancies (MMD), which are distances between embeddings of 
distributions in reproducing kernel Hilbert spaces (RKHS).
This equivalence immediately relates energy statistics to
kernel methods often used in machine learning, and form the basis 
of our approach.

@@

Although clustering from energy statistics, in Euclidean spaces,
was considered in \cite{Kgroups}, 
the precise optimization problem behind this approach
remains obscure, as well as the connection with kernel methods.
The main theoretical contribution of this paper is to fill this gap.
Since the statistical potential energy is minimum when
distributions are equal, the principle behind clustering is to maximize 
the statistical energy,  enforcing probability distributions associated to 
each cluster to be different from one another. We provide a precise 
mathematical formulation to this statement, leading to a quadratically 
constrained quadratic program (QCQP) in the associated RKHS. This immediately
establishes the connection between energy statistics based clustering,
or \emph{energy clustering} for short, with kernel methods. Moreover,
our formulation holds beyond the Euclidean case, more precisely, it holds
for general semimetric spaces of negative type.
We also show that such QCQP is equivalent to kernel $k$-means 
optimization problem when the kernel is fixed. However, energy statistics
is able to fix a family of kernels. Therefore, one can see 
kernel $k$-means as a consequence of energy statistics theory.
The equivalence between kernel $k$-means, spectral clustering, and graph 
partitioning problems is well-known \cite{Dhillon,Dhillon2}. We further
demonstrate 
how these relations arise from a weighted version of energy statistics.

Our main 
algorithmic contribution is to use Hartigan's method \cite{Hartigan} to 
find local solutions of the above mentioned QCQP, which is NP-hard in general.
Hartigan's method was also used in \cite{Kgroups}, but without
any connection to kernels. More importantly, the
advantages
of Hartigan's over Lloyd's method was already demonstrated 
in some simple settings
\cite{Telgarsky,Slonin}, but apparently this method did not receive 
the deserved attention. To the 
best of our knowledge, Hartigan's method was not previously 
employed together with kernel methods. 
We provide a fully kernel based Hartigan's algorithm for clustering,
where the kernel is fixed by energy statistics. 
We make clear the advantages of this proposal versus 
Lloyd's method, which kernel $k$-means algorithm 
is based upon and will also be used 
to solve our QCQP. We show that both algorithms  have the same
time complexity, but Hartigan's method in kernel spaces offer several
advantages since it is able to escape local minima of Lloyd's method. 
Furthermore, in the examples considered in this paper, the proposed method
also provides superior performance compared spectral clustering, which
is more expensive and in fact solves a relaxed version of our QCQP.

Our numerical results provide compelling evidence that 
Hartigan's method applied to energy clustering 
is more accurate and robust than kernel $k$-means algorithm.
Furthermore, our experiments illustrate 
the flexibility of energy clustering, 
showing that it is able to perform accurately on data coming from 
very different distributions, contrary to $k$-means and GMM, for instance.
More specifically, the proposed method performs 
closely to $k$-means and GMM on normally distributed data, however,
it is significantly better on data that 
is not normally distributed. 
Its superiority in high dimensions
is striking, being more accurate than $k$-means and GMM 
even on Gaussian settings.

%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.
% However, the Computer Society has been known to put floats at the bottom.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].
